An investigation into the contentious phylogenetic results of the *Mixta* species

### 2021-04-19 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

# Results
Twelve and twenty-three unique models were used for nucleotide (Table 2) and amino acid sequences (Table 3) were needed to create phylogenetic gene trees.

```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
PM_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table2_PmodelsNT.csv", stringsAsFactors = FALSE)

kable(PM_NT, caption = "Table 2. According to model testing and the lowest BIC for nucleotide sequences, the number of genes required each phylogenetic tree model.")
```

```{r Table3, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
PM_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_PmodelsAA.csv", stringsAsFactors = FALSE)

kable(PM_AA, caption = "Table 3. According to model testing and the lowest BIC for amino acid sequences, the number of genes required each phylogenetic tree model.")
```

After the two filters, 799 of 954 homologous genes remained. The genes included those encoding rRNA sequences, enzymes, transporters, and housekeeping genes and were scattered across the *M. calida* and *M. gaviniae* genomes.

## Best Model vs Best Available Model
In MEGAX [@Kumar2018], the user can either calculate genetic distance or estimate a phylogenetic tree. However, although model testing accesses all possible models available in the software, all models are not available for genetic distance analysis. The GTR (General Time Reversible) and HKY (Hasegawa-Kishino-Yano) models and the +I (invariant sites) parameter are available only for phylogenetic tree analysis.

```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1. The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange). Models are for nucleotide sequences.", fig.width=6.5}
models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE)

ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2",
                    labels = c("Phylogenetic", "Genetic Distance")) +
  labs(y = "Percent of Genes \n Requiring Each Model",
       fill = "Analysis") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels
        legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner
        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
        legend.text.align = 0,                                                            # Aligns legend text to the left
        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
        text = element_text(size = 12,  family = "Times New Roman"))
```

```{r Figure2, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 2. The percentage of genes (n = 799) most closely related to any of the eight non-*Mixta* species. Closest relative was inferred using shortest genetic distance. Blue bars represent the nucleotide sequences and green bars represent the amino acid sequences. The lighter of each colour represents *M. calida* and the darker *M. gaviniae*.", fig.width=6.5}
number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv",
                             stringsAsFactors = FALSE) %>%
  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis",
                                              "E. cloacae", "P. syringae"), ordered = TRUE),
         Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE))

ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Paired",
                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")),
                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")),
                               expression(paste("AA and ", italic("M. calida"), sep = "")),
                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
  labs(y = "Percent of Genes",
       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
        legend.text.align = 0,                                                            # Aligns legend text to the left
        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
        text = element_text(size = 12,  family = "Times New Roman"))
```

```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2. The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended (best) model from model testing vs the restricted model (best available) options available for genetic distance analysis in MEGAX for nucleotide sequences. The species with the shortest genetic distance from both of the *Mixta* species is considered the closest relative. Evolutionary patterns may vary by gene; therefore, it is appropriate to conduct model testing and use the model that explains the most variation. Nevertheless, this model may not be available in software programs or analysis methods within a software program. Pink bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the recommended model. Orange bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the best available genetic distance model analysis in MEGAX. GTR and HKY models and invariant sites parameter were not available for this analysis. The lighter shades represent *M. calida*, and the darker shades represent *M. gaviniae*.", fig.width=6.5}
number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeBestvsAvailableModelNT.csv",
                             stringsAsFactors = FALSE) %>%
  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis",
                                              "E. cloacae", "P. syringae"), ordered = TRUE),
         Type = factor(Type, levels = c("PD_calida", "PD_gaviniae", "GD_calida", "GD_gaviniae"), ordered = TRUE))

ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = c("#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00"),
                    labels = c(expression(paste("Best Model and ", italic("M. calida"), sep = "")),
                               expression(paste("Best Model and ", italic("M. gaviniae"), sep = "")),
                               expression(paste("Best Available Model and ", italic("M. calida"), sep = "")),
                               expression(paste("Best Available Model and ", italic("M. gaviniae"), sep = "")))) +
  labs(y = "Percent of Genes",
       fill = expression(paste("Model and ", italic("Mixta"), " species", sep = ""))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
        legend.text.align = 0,                                                            # Aligns legend text to the left
        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
        text = element_text(size = 12,  family = "Times New Roman"))
```

```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3. Closest relatives to *M. calida* genes for nucleotide sequences around the *M. calida* genome. Length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance. The *M. calida* genome has 4084 genes. No distinct patterns emerged; however, genes most closely related to *P. septica* are the most common and occur throughout the genome. Therefore, *P. septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5}
FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
  mutate(Closest_Relative = factor(Closest_Relative,
                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos",
                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))

ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
  coord_polar() +
  scale_fill_brewer(palette = "Paired",
                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
                               "P. syringae"),
                    na.value = "#ff7f00") +
  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  labs(fill = "Closest Relative") +
  theme_bw() +
  theme(legend.text = element_text(face = "italic"),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        text = element_text(size = 12,  family = "Times New Roman"))
```

```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4. Closest relatives to *M. calida* genes for amino acid sequences around the *M. calida* genome. Length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance. The *M. calida* genome has 4084 genes. No distinct patterns emerged; however, genes most closely related to *P. septica* are the most common and occur throughout the genome. Therefore, *P. septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5}
FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>%
  mutate(Closest_Relative = factor(Closest_Relative,
                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos",
                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))

ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
  coord_polar() +
  scale_fill_brewer(palette = "Paired",
                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
                               "P. syringae"),
                    na.value = "#ff7f00") +
  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  labs(fill = "Closest Relative") +
  theme_bw() +
  theme(legend.text = element_text(face = "italic"),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        text = element_text(size = 12,  family = "Times New Roman"))
```

```{r Figure5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 5. The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using percent site identities. The species with the highest percent identity to each of the *Mixta* species is considered the most similar taxon for a single gene. Blue bars represent the nucleotide sequences' results; the light blue bars represent *M. calida* sequences, and the dark blue bars represent *M. gaviniae* sequences. Green bars represent the amino acid sequences' results; the light green bars represent *M. calida* sequences, and the dark green bars represent *M. gaviniae* sequences.", fig.width=6.5}
number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure5_FirstRelativeMixtaNTAAIdentity.csv",
                             stringsAsFactors = FALSE) %>%
  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis",
                                              "E. cloacae", "P. syringae"), ordered = TRUE),
         Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE))

ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Paired",
                    labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")),
                               expression(paste("NT and ", italic("M. gaviniae"), sep = "")),
                               expression(paste("AA and ", italic("M. calida"), sep = "")),
                               expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) +
  labs(y = "Percent of Genes",
       fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
        legend.text.align = 0,                                                            # Aligns legend text to the left
        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
        text = element_text(size = 12,  family = "Times New Roman"))
```

```{r Figure6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 6. The most similar taxa to *M. calida* genes for nucleotide sequences around its genome. The most similar taxon is that which sequence is most identical to the *M. calida* sequence. Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5}
FI_calida_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure6_CircularIdentities_NT.csv",
                         stringsAsFactors = FALSE) %>%
  mutate(Closest_Relative = factor(Closest_Relative,
                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos",
                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))

ggplot(FI_calida_NT, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
  coord_polar() +
  scale_fill_brewer(palette = "Paired",
                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
                               "P. syringae")) +
  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  labs(fill = "Most Similar Taxa") +
  theme_bw() +
  theme(legend.text = element_text(face = "italic"),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        text = element_text(size = 12,  family = "Times New Roman"))
```

```{r Figure7, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 7. The most similar taxa to *M. calida* genes for amino acid sequences around its genome. The most similar taxon is that which sequence is most identical to the *M. calida* sequence. Length and colour of the bars represent the different species to which the gene is most similar.", fig.width=6.5, fig.height=5}
FI_calida_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure7_CircularIdentities_AA.csv",
                         stringsAsFactors = FALSE) %>%
  mutate(Closest_Relative = factor(Closest_Relative,
                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos",
                                              "T_saanichensis", "E_cloacae", "P_syringiae"), ordered = TRUE))

ggplot(FI_calida_AA, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 15) +
  coord_polar() +
  scale_fill_brewer(palette = "Paired",
                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
                               "P. syringae")) +
  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  labs(fill = "Most Similar Taxa") +
  theme_bw() +
  theme(legend.text = element_text(face = "italic"),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        text = element_text(size = 12,  family = "Times New Roman"))
```

```{r Table5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE}
stats_on_all <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/FigureO_AllStatsCalida.csv", stringsAsFactors = FALSE)

same_spp <- data.frame(table(stats_on_all$Same_Species))
same_spp <- same_spp[which(same_spp$Var1 != "Nope"), ]

same_spp <- rbind(same_spp, data.frame(Var1 = "T_ptyseos", Freq = 0))

colnames(same_spp) <- c("Species", "Number of Genes")

>>>>

# Results
```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE}
models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table2_BestAvailableModels.csv", stringsAsFactors = FALSE)

colnames(models)[2:3] <- c("Best Model", "Best Available Model")

kable(models, 
      caption = "Table 2. The number of genes that required each model according to the best model with the lowest BIC and the best available model for genetic distance analysis.")
```

```{r Figure1, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 1. The percentage of genes (n = 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended model from model testing versus the best available) model for genetic distance analysis in MEGAX. The sequences were nucleotide sequences. The species with the shortest genetic distance from the *Mixta* species is considered to be the closest relative. The pink and orange bars represent the percentage of *Mixta* genes that are most closely related to any of the non-*Mixta* species when using the recommended model and the best available model, respectively. The lighter shades represent *M. calida* and the darker shades represent *M. gavinaie*.", fig.width=6.5}
number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_FirstRelativeBestvsAvailableModelNT.csv", 
                             stringsAsFactors = FALSE) %>%
  mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", 
                                              "E. cloacae", "P. syringae"), ordered = TRUE),
         Type = factor(Type, levels = c("PD_calida", "PD_gaviniae", "GD_calida", "GD_gaviniae"), ordered = TRUE))

ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = c("#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00"),
                    labels = c(expression(paste("Recommended Model and ", italic("M. calida"), sep = "")), 
                               expression(paste("Recommended Model and ", italic("M. gaviniae"), sep = "")), 
                               expression(paste("Best Available Model and ", italic("M. calida"), sep = "")), 
                               expression(paste("Best Available Model and ", italic("M. gaviniae"), sep = "")))) +
  labs(y = "Percent of Genes", 
       fill = expression(paste("Model and ", italic("Mixta"), " species", sep = ""))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels
        legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner
        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
        legend.text.align = 0,                                                            # Aligns legend text to the left
        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
        text = element_text(size = 12,  family = "Times New Roman"))
```

<!-- # Old Results -->
<!-- ```{r Table2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=TRUE} -->
<!-- PM_NT <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table2_PmodelsNT.csv", stringsAsFactors = FALSE) -->

<!-- kable(PM_NT, caption = "Table 2. According to model testing and the lowest BIC for nucleotide sequences, the number of genes required each phylogenetic tree model.") -->
<!-- ``` -->

<!-- ```{r Table3, echo = FALSE, message = FALSE, warning=FALSE, fig.cap=TRUE} -->
<!-- PM_AA <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Table3_PmodelsAA.csv", stringsAsFactors = FALSE) -->

<!-- kable(PM_AA, caption = "Table 3. According to model testing and the lowest BIC for amino acid sequences, the number of genes required each phylogenetic tree model.") -->
<!-- ``` -->

<!-- ```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1. The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange). Models are for nucleotide sequences.", fig.width=6.5} -->
<!-- models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE) -->

<!-- ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
<!--   scale_fill_brewer(palette = "Dark2", -->
<!--                     labels = c("Phylogenetic", "Genetic Distance")) + -->
<!--   labs(y = "Percent of Genes \n Requiring Each Model", -->
<!--        fill = "Analysis") + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels -->
<!--         legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner -->
<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
<!-- ``` -->

<!-- ```{r Figure2, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 2. The percentage of genes (n = 799) most closely related to any of the eight non-*Mixta* species. Closest relative was inferred using shortest genetic distance. Blue bars represent the nucleotide sequences and green bars represent the amino acid sequences. The lighter of each colour represents *M. calida* and the darker *M. gaviniae*.", fig.width=6.5} -->
<!-- number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeNTAA.csv", -->
<!--                              stringsAsFactors = FALSE) %>% -->
<!--   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", -->
<!--                                               "E. cloacae", "P. syringae"), ordered = TRUE), -->
<!--          Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE)) -->

<!-- ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
<!--   scale_fill_brewer(palette = "Paired", -->
<!--                     labels = c(expression(paste("NT and ", italic("M. calida"), sep = "")), -->
<!--                                expression(paste("NT and ", italic("M. gaviniae"), sep = "")), -->
<!--                                expression(paste("AA and ", italic("M. calida"), sep = "")), -->
<!--                                expression(paste("AA and ", italic("M. gaviniae"), sep = "")))) + -->
<!--   labs(y = "Percent of Genes", -->
<!--        fill = expression(paste("Sequence type and ", italic("Mixta"), " species", sep = ""))) + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels -->
<!--         legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner -->
<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
<!-- ``` -->

<!-- ```{r Figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 2. The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species when using the recommended (best) model from model testing vs the restricted model (best available) options available for genetic distance analysis in MEGAX for nucleotide sequences. The species with the shortest genetic distance from both of the *Mixta* species is considered the closest relative. Evolutionary patterns may vary by gene; therefore, it is appropriate to conduct model testing and use the model that explains the most variation. Nevertheless, this model may not be available in software programs or analysis methods within a software program. Pink bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the recommended model. Orange bars represent the percentage of *Mixta* genes most closely related to non-*Mixta* species when using the best available genetic distance model analysis in MEGAX. GTR and HKY models and invariant sites parameter were not available for this analysis. The lighter shades represent *M. calida*, and the darker shades represent *M. gaviniae*.", fig.width=6.5} -->
<!-- number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure2_FirstRelativeBestvsAvailableModelNT.csv", -->
<!--                              stringsAsFactors = FALSE) %>% -->
<!--   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", -->
<!--                                               "E. cloacae", "P. syringae"), ordered = TRUE), -->
<!--          Type = factor(Type, levels = c("PD_calida", "PD_gaviniae", "GD_calida", "GD_gaviniae"), ordered = TRUE)) -->

<!-- ggplot(data = number_relatives, aes(x = Species, y = Percent, fill = Type)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
<!--   scale_fill_manual(values = c("#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00"), -->
<!--                     labels = c(expression(paste("Best Model and ", italic("M. calida"), sep = "")), -->
<!--                                expression(paste("Best Model and ", italic("M. gaviniae"), sep = "")), -->
<!--                                expression(paste("Best Available Model and ", italic("M. calida"), sep = "")), -->
<!--                                expression(paste("Best Available Model and ", italic("M. gaviniae"), sep = "")))) + -->
<!--   labs(y = "Percent of Genes", -->
<!--        fill = expression(paste("Model and ", italic("Mixta"), " species", sep = ""))) + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"),               # Rotates and italicizes x axis labels -->
<!--         legend.position = c(0.95, 0.95),                                                  # Moves legend inside the plot and in the top-right corner -->
<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
<!-- ``` -->

<!-- ```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3. Closest relatives to *M. calida* genes for nucleotide sequences around the *M. calida* genome. Length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance. The *M. calida* genome has 4084 genes. No distinct patterns emerged; however, genes most closely related to *P. septica* are the most common and occur throughout the genome. Therefore, *P. septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5} -->
<!-- FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>% -->
<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
<!--                                               "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE)) -->

<!-- ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
<!--   coord_polar() + -->
<!--   scale_fill_brewer(palette = "Paired", -->
<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
<!--                                "P. syringae"), -->
<!--                     na.value = "#ff7f00") + -->
<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
<!--   labs(fill = "Closest Relative") + -->
<!--   theme_bw() + -->
<!--   theme(legend.text = element_text(face = "italic"), -->
<!--         axis.title = element_blank(), -->
<!--         axis.text.y = element_blank(), -->
<!--         axis.ticks.y = element_blank(), -->
<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
<!-- ``` -->

<!-- ```{r Figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 4. Closest relatives to *M. calida* genes for amino acid sequences around the *M. calida* genome. Length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance. The *M. calida* genome has 4084 genes. No distinct patterns emerged; however, genes most closely related to *P. septica* are the most common and occur throughout the genome. Therefore, *P. septica* and *Mixta* are likely to have a recent common ancestor.", fig.width=6.5, fig.height=5} -->
<!-- FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>% -->
<!--   mutate(Closest_Relative = factor(Closest_Relative, -->
<!--                                    levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", -->
<!--                                               "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE)) -->

<!-- ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(), width = 15) + -->
<!--   coord_polar() + -->
<!--   scale_fill_brewer(palette = "Paired", -->
<!--                     labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae", -->
<!--                                "P. syringae"), -->
<!--                     na.value = "#ff7f00") + -->
<!--   scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) + -->
<!--   labs(fill = "Closest Relative") + -->
<!--   theme_bw() + -->
<!--   theme(legend.text = element_text(face = "italic"), -->
<!--         axis.title = element_blank(), -->
<!--         axis.text.y = element_blank(), -->
<!--         axis.ticks.y = element_blank(), -->
<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
<!-- ``` -->

<!-- ```{r Figure5, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 5. The percentage of genes (out of 799) that were most closely related to any of the eight non-*Mixta* species using percent site identities. The species with the highest percent identity to each of the *Mixta* species is considered the most similar taxon for a single gene. Blue bars represent the nucleotide sequences' results; the light blue bars represent *M. calida* sequences, and the dark blue bars represent *M. gaviniae* sequences. Green bars represent the amino acid sequences' results; the light green bars represent *M. calida* sequences, and the dark green bars represent *M. gaviniae* sequences.", fig.width=6.5} -->
<!-- number_relatives <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure5_FirstRelativeMixtaNTAAIdentity.csv", -->
<!--                              stringsAsFactors = FALSE) %>% -->
<!--   mutate(Species = factor(Species, levels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", -->
<!--                                               "E. cloacae", "P. syringae"), ordered = TRUE), -->
<!--          Type = factor(Type, levels = c("calida_NT", "gaviniae_NT", "calida_AA", "gaviniae_AA"), ordered = TRUE)) -->
